Eagle: multi-locus association mapping on a genome-wide scale made routine

### 2019-07-26 BioinformaticsRevison/bioinformaticspaper.tex
Working of stability selection section git push

 Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications. 
 It is used here to assign statistical significance to the regression coefficient estimates of
 LMM-Lasso, glmnet, and bigRR.   LMM-Lasso, glmnet, and bigRR are regularisation 
 methods. As such,  the size of a SNP effect in the model is estimated but not its significance.  
 Stability selection was chosen over permutation and other sampling procedures because of its reduced computational cost.

The Stability selection procedure for LMM-Lasso and glmnet is as follows. Remembering that for each simulation scenario we generate 100 
replicates of data under that scenario, we begin by choosing one of these replicates at random. 
We then analyse these data with LMM-Lasso (glmnet) with different values of its  regularisation parameter. 
Our goal is to find the parameter value that  yields 20 to 30 non-zero SNP effects. 
This can be done via a simple binary search approach. 
 Then, for each replicate, we subsampled repeatedly, 100 subsets of size $n/2$.  A replicate is subsampled without replacement.  
 These subsets are then analysed with LMM-Lasso (glmnet) with its regularisation parameter fixed to the value found previously. 
A binary vector, of length the number of SNP, is recorded as the result. A one (zero) means the SNP had a non-zero (zero) effect size. 
Calculating a SNP's statistical significance is now a simple task. We take the vector sum of the binary vectors over all 100 subsets. 
This vector sum will contain elements in the range of 0 to 100. By dividing each element in this vector sum by the number of 
subsets upon which the sum was calculated, we obtain empirical probabilities.  
 These 
probabilities measure the strength of evidence for the SNPs to be in association with the trait.

By taking the vector sum over all 100 subsets and dividing each element in the sum by 100, we obtain the probability 
estimates of the significance of the SNPs.

A probability estimate of the significance of a SNP is now just a simple vector sum, over the 100 subsets,  of the binary vectors

By taking the vector sum over all 100 subsets, for a particular replicate, and dividing each element sum by 100, we obtain a probability 
estimate of significance.

Now, for a replicate,  to calculate the statistical significance of the coefficient estimates obtained with LMM-Lasso (glmnet),

estimate the statistical significance of a SNP, for a particular replicate, is obtained by taking the

Now, for a particular replicate, we count the number of times LMM-Lasso (glmnet) gave a non-zero

Now, for a particular replicate which was analysed with LMM-Lasso (glmnet), 
we count the number of times

we calculate the statistical significance of a SNP by counting the number of times its 
coefficient was non-zero in the LMM-Lasso (glmnet) analyses of the 100 subsets of data.  A probability estimate of significance is obtained 
by dividing this count by the number of subsetted analyses, which was 100.  Note that it is not 
necessary to tune the regularisation   parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. Also, 
we have the luxury of knowing the genetic conditions 
under which the simulated trait data are generated. we know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
 However, stability selection is a subsampling procedure that is robust to parameter misspecification \citep{meinshausen2010stability}. 
 The regularisation parameter can be tuned to any reasonable number of 
 non-zero effects.

Then, for the LMM-Lasso (glmnet) analysis of a given replicate, to estimate the statistical significance of a SNP effect, we count 
the number of times the SNP effect has non-zero effect size across the 100 subset LMM-Lasso (glmnet) analyses.  This number is then 
divided by the number of subsets (which is 100 for the replicate)

 to estimate the statistical significance of the SNP effects for a given replicate, the number of times

For a replicate, a (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
number of times a SNP has a non-zero effect size over all the subsets divided by the number of subsets (which was 100).

Here, the link between 
trait and genotype is preserved, with only the sample size being reduced.

\textcolor{red}{
To obtain significance estimates via stability selection, we do the following. For LMM-Lasso and glmnet, for a given scenario, 
we begin by performing a preliminary analysis of a randomly chosen  replicate (i.e. trait and SNP genotype data)  from the 
 100 available replicates for that scenario.  Our goal is to find a value of the regularisation parameter used by LMM-Lasso and glmnet 
 that yields 20 to 30 non-zero SNP effects. It is not necessary to tune the regularisation 
  parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. 
  Since we have the luxury of knowing the genetic conditions 
under which the simulated trait data are generated, we know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
 However, stability selection is a subsampling procedure that is robust to misspecification of parameters  \citep{meinshausen2010stability}. 
 The regularisation parameter can be tuned to any reasonable number of 
 non-zero effects.  Once we have a value for the regularisation parameter,

 We subsample repeatedly, without replacement, from a replicate. We draw 100 subsets of size $n/2$.  
Each subset is then analysed with LMM-Lasso and glmnet, with their regularisation parameter set to the value found in the preliminary analysis. 
A (probability) estimate of the statistical significance of a SNP effect is obtained by counting the 
number of times the SNP have a non-zero effect size over all the replicates divided by the number of subsets (which was 100). }

For bigRR, we modify our stability selection procedure slightly. 
There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
number of times the SNP is recorded divided by the number of replicates.

>>>>

 Stability selection \citep{meinshausen2010stability}  is a subsampling strategy with a range of applications. 
 It is used here to estimate, empirically, the statistical significance of the results from LMM-Lasso, glment, and bigRR 
 of the simulated data. These three regularisation methods give the effect sizes of the SNPs, but not their significance, as their 
 results.   Stability selection was chosen over permutation and other sampling procedures because of its reduced computational cost.

The Stability selection procedure for LMM-Lasso and glmnet is as follows. 
For a particular scenario, we begin by finding the value of the regularisation parameter that yields 20 to 30 non-zero SNP effects. 
We know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect from the analysis of a replicate in the simulation study.  The regularisation parameter though could have been tuned to give any reasonable number of non-zero SNP effects. 
 This tuning was done for each of the six scenarios 
 but only for a single replicate within a scenario. It is not necessary to tune the regularisation parameter on every replicate 
 when replicates are generated under the same (trait, sample size, and number of SNP) conditions within a scenario.

We did this for each  scenarios but not for every replicate within a scenario. It is unnecessary to set the regularisation parameter 
for every replicate when replicates within a scenario are generated under the same conditions. 
For a given scenario, we chose a replicate at random.  We then used a binary search procedure to set the regularisation parameter. 
Then, for a replicate, we subsample repeatedly, 100 subsets of size $n/2$. These subsamples are drawn without replacement.  
Here, the matching of trait to genotype is preserved in the subsamples. A subsample differs to the replicate in size only.

Remembering that for each simulation scenario we generate 100 
replicates of data under that scenario, we begin by choosing one of these replicates at random. 
We then analyse this replicate with LMM-Lasso (glmnet) with different values of its  regularisation parameter. 
Our goal is to find the parameter value that  yields 20 to 30 non-zero SNP effects. 
This can be done via a simple binary search approach. 
 Then, for each replicate, we subsampled repeatedly, 100 subsets of size $n/2$.  A replicate is subsampled without replacement.  
 These subsets are then analysed with LMM-Lasso (glmnet) with its regularisation parameter fixed to the value found previously. 
A binary vector, of length the number of SNP, is recorded as the result. 
A one (zero) means the SNP had a non-zero (zero) effect size. 
Calculating a SNP's statistical significance is now a simple task. We take the vector sum of the binary vectors over all 100 subsets. 
This vector sum will contain elements in the range of 0 to 100. By dividing each element in this vector sum by the number of 
subsets upon which the sum was calculated, we obtain empirical probabilities.  
 These 
probabilities measure the strength of evidence for the SNPs to be in association with the trait. 
 Note that it is not 
necessary to tune the regularisation   parameter on every replicate within a scenario because the replicates are generated under the same conditions within a scenario. Also, 
we have the luxury of knowing the genetic conditions 
under which the simulated trait data are generated. we know that 20 to 30 SNP-trait associations is a reasonable number of findings to expect. 
 However, stability selection is a subsampling procedure that is robust to parameter misspecification \citep{meinshausen2010stability}. 
 The regularisation parameter can be tuned to any reasonable number of 
 non-zero effects.

For bigRR, we modify our stability selection procedure slightly.  Where LMM-Lasso and glmnet shrink non-important effects to zero, 
bigRR is a ridge regression method where all SNP are given an effect size.  XXXXX expand on this. ...

There is no need to tune the regularisation parameter for bigRR as an optimal value is found as part of its analysis. 
As described above, we draw 100 replicate subsets of size $n/2$ and analyse these data with bigRR. We then order the 
SNPs according to the absolute size of their SNP effects and record the top 20 SNPs. 
A (probability)  estimate of the statistical significance of a SNP effect is then obtained by counting the 
number of times the SNP is recorded divided by the number of replicates.
